\centerline{Steve Hsu\hfill extra credit 5}
\item{a.}

We will prove the claim by contraposition, that is,
if $\sum a_n / (1 + a_n)$ converges, then $\sum a_n$ converges.
Assume that $\sum a_n / (1 + a_n)$ converges.
We will first show that eventually $a_n < 1$.
Since $\sum a_n / (1 + a_n)$ converges, it must be Cauchy.
Taking $\epsilon = 1/2$, there is a natural number $N$
such that $a_n / (1 + a_n) < 1/2$ for all $n \ge N$.
Rearranging the inequality, we have that
$2a_n < 1 + a_n$ and therefore $a_n < 1$ for all $n \ge N$.
Therefore, eventually $a_n < 1$.
If $a_n < 1$, then $1 + a_n < 1 + 1 = 2$ and
$1 / (1 + a_n) > 1/2$.
Therefore, eventually $a_n / (1 + a_n) > a_n / 2$.
Since eventually $0 < a_n / 2 < a_n / (1 + a_n)$
and $\sum a_n / (1 + a_n)$ converges,
by the comparison test, $\sum a_n / 2$ converges as well.
By the algebraic limit theorem for series,
it follows that $\sum a_n$ converges, as desired.
\bigskip
\item{b.} let $a_n = 1/n$

We know that $\sum 1/n$, which is the harmonic series, diverges.

Since $(1/n) / (1 + n(1/n)) = (1/n) / (1 + 1) = 1/(2n)$,
and $\sum 1/(2n)$ is just half of the harmonic series,
it diverges.
\bigskip
\item{c.}
$$a_n = \cases{
1/\sqrt n,&if $n = k^2$ for some natural number $k$\cr
1/n^2,&otherwise\cr
}$$
Notice that $a_1 + a_4 + a_9 + \cdots = 1 + 1/2 + 1/3 + \cdots$
is the harmonic series, which diverges,
so clearly, $\sum a_n$, which has only positive terms,
must diverge as well.

If $n = k^2$ for some natural number $k$, then
$a_n / (1 + na_n) = (1/k) / (1 + n(1/k)) = (1/k) / (1 + k^2/k) = 1 / (k + k^2)$,
so the sum over perfect squares of $a_n / (1 + na_n)$ converges.
For all other values of $n$,
$a_n / (1 + na_n) = (1/n^2) / (1 + n(1/n^2)) = 1 / (n^2 + n)$,
so the partial sums over numbers not perfect squares
are at most those of $\sum 1/(n^2 + n)$ and therefore converge.
Since $a_n$ is positive for all $n$,
the partial sums are monotonically increasing and bounded above by
$(\sum 1 / (k + k^2)) + (\sum 1 / (n^2 + n))$,
so by the monotone convergence theorem, the sequence of partial sums converges.
\bye
